A highly-available move operation for replicated trees

### 2019-06-20 paper/move-op.tex
paper: editing, removing lines with a few words
Some small changes to experiment write-up and code-gen sections.

\subsection{Extracting efficient algorithms from HOL}

@@

Write $('a::\{hashable\}, 'b)\ \isa{hm}$ for the type of hash-maps with key type $\isa{'a}$ and binding type $\isa{'b}$.
Note that keys have an explicit $\isa{hashable}$ constraint, meaning that this hash-map type may only be used with kay types with an associated hashing function.
With this type, we represent trees by the type $(\isa{'n}::\{hashable\}, \isa{'m} \times \isa{'n})\ \isa{hm}$, mapping from a \emph{child} node to its \emph{parent} node and associated metadata, if any such mapping exists.

@@

Intuitively, this states that if both algorithms are applied to the same set of operations then they produce identical logs and also produce trees that contain the same set of parent-metadata-child records---i.e. the two trees are extensionally equivalent, despite having very different concrete representations.
Moreover, we have the following corollaries of the theorems \isa{apply\_ops\_commutes} and \isa{apply\_ops\_acyclic}, respectively, further establishing the correctness of our efficient algorithms:\footnote{The ``efficient'' analogue of \isa{apply\_ops\_unique\_parent} is trivially true, as the output tree is stored as a hash-map, and therefore too trivial to state as a theorem.}:

@@

To evaluate the viability of our algorithm in practice, we used Isabelle/HOL's code generation facility to extract executable Scala code from our formally verified definitions.
We wrapped this implementation in a network service using the Akka~\cite{Akka} actor framework, using gRPC~\cite{gRPC} for network communication.
We deployed three replicas of this service on Amazon EC2 t2.small instances in different regions worldwide: one in Northern California (region us-west-1), one in Ireland (eu-west-1), and one in Singapore (ap-southeast-1).
The network delays for communication between these regions are substantial; see Table~\ref{tab:rpc-times}.

>>>>

\subsection{Extracting efficient algorithms from HOL} \label{subsect.extracting}

@@

Write $('a::\{hashable\}, 'b)\ \isa{hm}$ for a hash-map type with key type $\isa{'a}$ and binding type $\isa{'b}$.
Note that keys have a $\isa{hashable}$ constraint, meaning that the type may only be used with key types with a defined hashing function.
With this type, we represent trees by the type $(\isa{'n}::\{hashable\}, \isa{'m} \times \isa{'n})\ \isa{hm}$, mapping from a \emph{child} node to its \emph{parent} node and associated metadata, if any such mapping exists.

@@

That is, if both algorithms are applied to the same set of operations then they produce identical logs and also produce trees that contain the same set of key-value bindings---i.e. the two trees are extensionally equivalent, despite having very different concrete representations.
Moreover, we have the following corollaries of the theorems \isa{apply\_ops\_commutes} and \isa{apply\_ops\_acyclic}, respectively, further establishing the correctness of our efficient algorithms:\footnote{The ``efficient'' analogue of \isa{apply\_ops\_unique\_parent} is trivially true, as the output tree is stored as a hash-map, and therefore too trivial to state as a theorem.}:

@@

We used Isabelle/HOL's code generation mechanism to extract executable Scala code from our HOL definitions, as discussed in Subsection~\ref{subsect.extracting}, and wrapped this implementation in a network service using the Akka~\cite{Akka} actor framework, using gRPC~\cite{gRPC} for network communication.
We deployed three replicas of this service on Amazon EC2 \texttt{t2.small} instances in different regions worldwide: one in Northern California (region \texttt{us-west-1}), one in Ireland (\texttt{eu-west-1}), and one in Singapore (\texttt{ap-southeast-1}).
The network delays for communication between these regions are substantial; see Table~\ref{tab:rpc-times}.
